I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-27 Paper/comppaths.tex
proof

The sketch of proof of Theorem \ref{th2} can be found in Section \ref{sec.proofs}.

\subsection{Sketch of Proofs}
\label{sec.proofs}

@@

%Intuitively, our first proof shows that if there are no compensation, then there are no correlations between nodes.
Consider a target neuron $z$.
In case there is no compensating path, we can assign a sign to each neuron $n$: 0
if all paths from $n$ to $z$ has weight 0, $+1$ if all paths have positive weights, and 
$-1$ if they have negative weights. Then we prove:

\begin{proposition}
	\label{prop.sign}
	There exist two inputs $\vx^\star,\vx^\sharp$ such that ${\vx^\star}$ maximizes the value of all positive neurons and minimizes the value of all negative neurons, and  ${\vx^\sharp}$ minimizes the value of all positive neurons and maximizes the value of all negative neurons.
\end{proposition}

@@

\smallskip

@@

The formal proofs are in appendix.

>>>>

The sketch of proof of Theorem \ref{th2} can be found in Section \ref{sec.proofs2}.

\subsection{Sketch of Proof for Theorem \ref{th1}}
\label{sec.proofs}

@@

%Intuitively, our first proof shows that if there are no compensation, then there are no correlations between nodes.
Consider a target neuron $z$.
In case there is no compensating path, we can assign a sign to each neuron $n$: 0
if all paths from $n$ to $z$ has weight 0, $+1$ if all paths have positive weights, and 
$-1$ if they have negative weights.

Using the concept of sign, we introduce input vectors $\vx^{*}, \vx^{\sharp}$:

\begin{definition}
We define the following two input vectors $\vx^{*}, \vx^{\sharp}$ in $\cal B$: 
	\begin{itemize}
		\item $\vx^{*}$ is the input vector defined in the following way:
		\begin {itemize}
		 \item $x^*_{a_i}=\max(a_i)$ if $S(a_i)\in \{0,1\}$, and
          \item $x^*_{a_i}=\min(a_i)$ otherwise, that is when $S(a_i)=-1$.
	\end{itemize}

		\item $\vx^{\sharp}$ is the input vector defined in the following way:
		\begin{itemize}
			\item $x^{\sharp}_{a_i}=\min(a_i)$ if $S(a_i)\in \{0,1\}$, and
			\item $x^{\sharp}_{a_i}=\max(a_i)$ otherwise, that is when $S(a_i)=-1$.
		\end{itemize}
	\end{itemize}
\end{definition}

We can then prove that $\vx^\star,\vx^\sharp$ maximize and minimize conjointly all the nodes in the DNN. Notice that it does {\em not} mean that there is no dependencies between nodes.
There can be correlations between nodes, just they do not affect the $\min$ and $\max$ bounds that can be reached in any node.

\begin{proposition}
	\label{prop.sign}
	${\vx^\star}$ maximizes the value of all positive neurons and minimizes the value of all negative neurons, and  
	${\vx^\sharp}$ minimizes the value of all positive neurons and maximizes the value of all negative neurons.
\end{proposition}

@@

\smallskip
The formal proof is in appendix A.

\subsection{Sketch of Proof for Theorem \ref{th2}}
\label{sec.proofs2}

@@

\smallskip
The formal proof is in appendix.

### 2024-01-27 Paper/proofs.tex
proof

We prove by induction from the input layer to higher layers that the upper bound of node $n$ in $D'$, is greater than or equal to the upper bound in $D$; and similarly for the lower bound. We use  $\Val^{D}_{x_J^0}(n)$ to denote the range of possible values for node $n$ in  model $D$ or $D'$ when $x_J=x_J^0$. 
This may be a single number or an interval. 
Denote by $\beta_{x_J^0}(m)$ (resp $\beta'_{x_J^0}(m)$) the upper bound computed by $\MILP_X$ 
for $m$ in $D$ (resp. $D'$) when the $J$ input is fixed to be ${x_J^0}$.

For any input $x_J=x_J^0$, we prove $\beta'_{x_J^0} \geq \beta_{x_J^0}$ by induction on $n$:

\begin{align*}
	\beta'_{x_J^0}(n) = & {bias}_n+\sum_{m\in B_{pure}} W_{mn} \beta'(\hat{m})
	+\sum_{m\notin B_{pure}} W_{mn} (\beta'_{x_J^0}(\hat{m}))\\
(\text{by induction})\geq & {bias}_n+\sum_{m\in B_{pure}} \max(\ReLU(W_{mn}m))+\sum_{m\notin B_{pure}} 
\UB(\ReLU(W_{mn}\Val^{D}_{x_J^0}(m)))\\
= & {bias}_n+\sum_{m\in B_{pure}} \UB(\mathrm{AppB}(W_{mn}))+\sum_{m\notin B_{pure}}  \UB(\ReLU(W_{mn}\Val^{D}_{x_J^0}(m)))\\
\geq & \UB(\Val^{D}_{x_J^0}(n)), \text{when } S(n)=1
\end{align*}

>>>>

We prove by induction from the input layer to higher layers that the upper bound of node $n$ in $D'$, is greater than or equal to the upper bound in $D$; and similarly for the lower bound. We use  $\Val^{D}_{x_J^0}(n)$ to denote the range of possible values for node $n$ in  model $D$ or $D'$ when $x_J=x_J^0$. 
This may be a single number or an interval. 
Denote by $\beta_{x_J^0}(m)$ (resp $\beta'_{x_J^0}(m)$) the upper bound computed by $\MILP_X$ 
for $m$ in $D$ (resp. $D'$) when the $J$ input is fixed to be ${x_J^0}$.
We define in the same way  $\alpha_{x_J^0}(m)$ and $\alpha'_{x_J^0}(m)$ for the minimal values.

For any input $x_J=x_J^0$, we prove $\beta'_{x_J^0}(n) \geq \beta_{x_J^0}(n)$ 
and
$\alpha'_{x_J^0}(n) \leq \alpha_{x_J^0}(n)$ 
by induction on $n$:

\begin{align*}
\beta'_{x_J^0}(n) = & {bias}_n+\sum_{m\in B_{pure},W_{mn}>0} W_{mn} \beta'(\hat{m})+\sum_{m\in B_{pure},W_{mn}<0} W_{mn} \alpha'(\hat{m})\\
+&\sum_{m\notin B_{pure},W_{mn}>0} W_{mn} \beta'_{x_J^0}(\hat{m})+\sum_{m\notin B_{pure},W_{mn}<0} W_{mn} \alpha'_{x_J^0}(\hat{m})\\
\\
(\text{by induction})\geq & {bias}_n+\sum_{m\in B_{pure},W_{mn}>0} W_{mn} \beta'(\hat{m})+\sum_{m\in B_{pure},W_{mn}<0} W_{mn} \alpha'(\hat{m})\\
+&\sum_{m\notin B_{pure},W_{mn}>0} W_{mn} \beta_{x_J^0}(\hat{m})+\sum_{m\notin B_{pure},W_{mn}<0} W_{mn} \alpha_{x_J^0}(\hat{m})\\
\\
(\text{Lemma \ref{lem:key1}})= & {bias}_n+\sum_{m\in B_{pure},W_{mn}>0} W_{mn} \beta(\hat{m})+\sum_{m\in B_{pure},W_{mn}<0} W_{mn} \alpha(\hat{m})\\
+&\sum_{m\notin B_{pure},W_{mn}>0} W_{mn} \beta_{x_J^0}(\hat{m})+\sum_{m\notin B_{pure},W_{mn}<0} W_{mn} \alpha_{x_J^0}(\hat{m})\\
\geq & \beta_{x_J^0}(n)
\end{align*}

The case for $\alpha$ is similar, proving the induction.
